A point-level trend model for the North American Breeding Bird Survey that corrects for detectability

### 2024-10-16 article/data-chapter-3.tex
renamed file and got rid of abstract for thesis version
remove conservation implications and move to own chapter

\section{Abstract}
\par TO DO

@@

\begin{table}[!ht]
	\caption{Parameters in the BBS models. Equation \ref{bbs} refers to the route-level model; Equation \ref{bbs-point} refers to the point-level model; Equation \ref{bbs-detect} refers to the point-level model with detectability offsets; and Equation \ref{bbs_varprop} refers to the point-level model with detectability and variance propagation.}
	\label{params_table}
		\begin{tabular}[t]{l>{\raggedright}p{0.5\linewidth}>{\raggedright\arraybackslash}p{0.3\linewidth}}
		\hline
		Parameter or Datum&Description&Model\\
		\hline
		$i$&Stratum index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$j$&Point/stop index&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$J$&Route index&Equation \ref{bbs}\\
		$t$&Year index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$o$&Observer index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$Y_{J,i,t,o}$&Total observed birds at route $J$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs}\\
		$Y_{j,i,t,o}$&Total observed birds at point $j$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\alpha_i$&Stratum-level intercept, modelled with an intrinsic conditional autoregressive model \citep{besag_bayesian_1991, smith_spatially_2023}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\delta_{i,t}$&Smoothed temporal component estimating population trajectory through time. Modelled with generalized additive model \citep{smith_north_2021}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\gamma_{i,t}$&Annual year-effect fluctuation term; departure from smoothed GAM&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\psi_J$&Route-level intercept&Equation \ref{bbs}\\
		$\psi_j$&Point-level intercept&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\omega_o$&Observer-level intercept&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\eta$&First-year observer effect&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$I(o,J,t)$&1 if observer is surveying route $J$ in year $t$ for the first time, 0 otherwise&Equation \ref{bbs}\\
		$I(o,j,t)$&1 if observer is surveying point $j$ in year $t$ for the first time, 0 otherwise&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$p_{j,i,t}$&Probability of availability at point $j$ in stratum $i$ in year $t$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$q_{j,i,t}$&Probability of perceptibility at point $j$ in stratum $i$ in year $t$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\kappa_{j,i,t}^{(p)}$&Design matrix for availability variance propagation; first derivatives of probability of availability \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\kappa_{j,i,t}^{(q)}$&Design matrix for perceptibility variance propagation; first derivatives of probability of perceptibility \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\zeta$&Effects vector for availability variance propagation \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\xi$&Effects vector for perceptibility variance propagation \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		\hline
	\end{tabular}
\end{table}%

>>>>

@@

\begin{table}[!ht]
	\caption{Parameters in the BBS models. Equation \ref{bbs} refers to the route-level model; Equation \ref{bbs-point} refers to the point-level model; Equation \ref{bbs-detect} refers to the point-level model with detectability offsets; and Equation \ref{bbs_varprop} refers to the point-level model with detectability and variance propagation.}
	\label{params_table}
		\begin{tabular}[t]{l>{\raggedright}p{0.5\linewidth}>{\raggedright\arraybackslash}p{0.3\linewidth}}
		\hline
		Parameter or Datum&Description&Model\\
		\hline
		$i$&Stratum index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$j$&Point/stop index&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$J$&Route index&Equation \ref{bbs}\\
		$t$&Year index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$o$&Observer index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$Y_{J,i,t,o}$&Total observed birds at route $J$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs}\\
		$Y_{j,i,t,o}$&Total observed birds at point $j$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\alpha_i$&Stratum-level intercept, modelled with an intrinsic conditional autoregressive model \citep{besag_bayesian_1991, smith_spatially_2023}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\delta_{i,t}$&Smoothed temporal component estimating population trajectory through time. Modelled with generalized additive model \citep{smith_north_2021}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\gamma_{i,t}$&Annual year-effect fluctuation term; departure from smoothed GAM&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\psi_J$&Route-level intercept&Equation \ref{bbs}\\
		$\psi_j$&Point-level intercept&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\omega_o$&Observer-level intercept&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\eta$&First-year observer effect&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$I(o,J,t)$&1 if observer is surveying route $J$ in year $t$ for the first time, 0 otherwise&Equation \ref{bbs}\\
		$I(o,j,t)$&1 if observer is surveying point $j$ in year $t$ for the first time, 0 otherwise&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$p_{j,i,t}$&Probability of availability at point $j$ in stratum $i$ in year $t$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$q_{j,i,t}$&Probability of perceptibility at point $j$ in stratum $i$ in year $t$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		\hline
	\end{tabular}
\end{table}

\begin{table}[!ht]
	\caption{Table \ref{params_table} continued}
	\begin{tabular}[t]{l>{\raggedright}p{0.5\linewidth}>{\raggedright\arraybackslash}p{0.3\linewidth}}
		\hline
		Parameter or Datum&Description&Model\\
		\hline
		$\kappa_{j,i,t}^{(p)}$&Design matrix for availability variance propagation; first derivatives of probability of availability \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\kappa_{j,i,t}^{(q)}$&Design matrix for perceptibility variance propagation; first derivatives of probability of perceptibility \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\zeta$&Effects vector for availability variance propagation \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\xi$&Effects vector for perceptibility variance propagation \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		\hline
	\end{tabular}
\end{table}
